A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2016-06-10 LensingPaper_draft_v4.01.tex
Tried to make the M/L section more definit.

\NEW{A fit to all nine available $v_\text{rms}$ data points forces the DM halo to be less massive. But with decreasing DM contribution in the inner regions the model approaches the mass-follows-light model in the center, that we already ruled out in Section \ref{sec:results_JAM_SB_MfL}. Allowing for free $\Upsilon_{*}(R')$, $\beta_z(R')$ and/or non-NFW $\rho_\text{halo}(r)$ profiles would over-fit the data again. Without further priors on the stellar $\Upsilon_{*}(R')$ and/or more data points, we can make only the following statement: A model with constant $\Upsilon_*$, moderately tangential \Wilma{[TO DO: Check]} $\beta_z$ and a spherical and massive DM component with high concentration can explain the central kinematics of J1331 very well, we have however ruled out that in this case the DM halo will follow a NFW halo at both small and large radii.}
%--------------------------------------------

@@

Even though investigation of the photometry of J1331 did not reveal a distinct blue core in J1331, we cannot fully exclude the possibility that J1331 has such a $\Upsilon_*$ gradient (see discussion in Sections \ref{sec:results_JAM_SB_MfL} and \ref{sec:results_JAM_SB_gradient}). \NEW{Stellar population analysis with spectra taken at different positions within the bulge could provide further information on the true $\Upsilon_*(R')$.}\OLD{In this case the restrictive assumption of a constant $\Upsilon_*$ in our JAM modelling could have lead to wrong conclusions about the DM distribution.}

@@

@@

When \citet{SWELLSI} assumed a Chabrier IMF, their result translates to $\Upsilon_\text{I,*}^\text{chab} = 2.5 \pm 0.6$ (see Table \ref{tab:previousresults}). In Section \ref{sec:results_JAM_SB_gradient} we created a dynamical model from only the surface brightness distribution and an increasing mass-to-light ratio profile without additional DM halo. We found that such a model would be perfectly consistent with the Einstein mass, predict a total $\Upsilon_\text{I,tot}(R'\sim0) = 2.53$---being consistent with the Chabrier IMF estimate by \citet{SWELLSI}---and rise quickly to $\Upsilon_\text{I,tot}(R'\gtrsim R_\text{ein}) \gtrsim 6$.

We also compare our results from Section \ref{sec:results_JAM_NFW} with the study by \citet{SWELLSV}. They found that the bulge of J1331 has an IMF \emph{more} bottom-heavy than the Salpeter IMF from fitting a NFW halo to (i) the Einstein mass and (ii) gas kinematics at larger radii $\gtrsim 8''$. In Section \ref{sec:results_JAM_NFW} we fitted a mass model with NFW halo to (i) the Einstein mass and (ii) stellar kinematics within $\sim 3.5''$. Our best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a \emph{less} bottom-heavy IMF than the Salpeter IMF. \cite{SWELLSV} found systematically lower NFW halo masses ($v_\text{circ,halo}(5'') \sim 120~\text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ,halo}(R'=5'') \sim 200~\text{km s}^{-1}$).
%--------------------------------------------

@@

@@

\NEW{Around $R'\sim7''$ the rotation curve and dispersion show some wiggles which lead to a spontaneous rise of the $v_\text{rms}$. As the spiral arms cross the major axis around this radius, we suspect they cause this disturbance of the axisymmetric kinematics. The DM halo should also start to dominate the kinematics at larger radii, whose total mass we expect more of the order of what \citep{Dutton paper} \Wilma{???} found---unless J1331 has a strong kinematic warp in its disk, which cannot be detected in long-slit kinematics only.}

>>>>

\NEW{A fit to all nine available $v_\text{rms}$ data points would force the DM halo to be less massive. But with decreasing DM contribution in the inner regions the model approaches the mass-follows-light model in the center, that we already ruled out in Section \ref{sec:results_JAM_SB_MfL}. Allowing for free $\Upsilon_{*}(R')$, $\beta_z(R')$ and/or non-NFW $\rho_\text{halo}(r)$ profiles would over-fit the data again. Without further priors on the stellar $\Upsilon_{*}(R')$ and/or more data points, we can make only the following statement: A model with constant $\Upsilon_*$, moderately tangential \Wilma{[TO DO: Check]} $\beta_z$ and a spherical and massive DM component with high concentration can explain the central kinematics of J1331 very well, we have however ruled out that in this case the DM halo will follow a NFW halo at both small and large radii.}
%--------------------------------------------

@@

Even though investigation of the photometry of J1331 did not reveal a distinct blue core in J1331, we cannot fully exclude the possibility that J1331 has such a $\Upsilon_*$ gradient (see discussion in Sections \ref{sec:results_JAM_SB_MfL} and \ref{sec:results_JAM_SB_gradient}). \NEW{Spatially resolved stellar population analysis based on integral-field spectroscopy of J1331's bulge could provide further information on the true $\Upsilon_*(R')$.}\OLD{In this case the restrictive assumption of a constant $\Upsilon_*$ in our JAM modelling could have lead to wrong conclusions about the DM distribution.}

@@

\NEW{In the following we discuss why we think---based on our results and previous analyses---that J1331's bulge has an IMF slightly less bottom-heavy than the Salpeter-like IMF.}

\NEW{\emph{(i) Indications for a Salpeter-like IMF in J1331's bulge:}}

@@

\NEW{\emph{(ii) Indications for and arguments against a Chabrier-like IMF in J1331's bulge:}}

When \citet{SWELLSI} assumed a Chabrier IMF, their result translates to $\Upsilon_\text{I,*}^\text{chab} = 2.5 \pm 0.6$ (see Table \ref{tab:previousresults}). In Section \ref{sec:results_JAM_SB_gradient} we created a dynamical model from only the surface brightness distribution and an increasing mass-to-light ratio profile without additional DM halo \NEW{and without velocity anisotropy}. We found that such a model would be perfectly consistent with the Einstein mass, predict a total $\Upsilon_\text{I,tot}(R'\sim0) = 2.53$---being consistent with the Chabrier IMF estimate by \citet{SWELLSI}---and rise quickly to $\Upsilon_\text{I,tot}(R'\gtrsim R_\text{ein}) \gtrsim 6$.

\NEW{This rise could be either due to a cuspy DM halo (see Section \Wilma{4.4???}) or a very strong gradient in $\Upsilon_\text{I,*}$ (see Section \Wilma{5.1???}). However, we rule out that a DM cusp is the sole reason because a cuspy and therefore overall massive DM halo does not match the kinematics in J1331's outer regions. We also rule out that a very strong $\Upsilon_\text{I,*}$ gradient is the only reason, because---as mentioned in Section \Wilma{???}---photometry did not reveal the clear existence of a blue population in the very center of J1331's bulge. Also, our modelling attempts allowing for velocity anisotropy (Sections \Wilma{4.3.2} and \Wilma{4.4}) suggest that we do need some moderate $\beta_z<0$ to explain the central $v_\text{rms}$ dip. And lastly---as laid out in the previous paragraph---J1331's central velocity dispersion suggests a more bottom-heavy IMF. Overall it is therefore not very likely that the central regions of J1331 have such a low $\Upsilon_\text{I,*}^\text{chab} \sim 2.5$.}

\NEW{\emph{(iii) Arguments against an IMF more bottom-heavy than the Salpeter IMF in J1331's bulge:}}

We also compare our results from Section \ref{sec:results_JAM_NFW} with the study by \citet{SWELLSV}. They found that the bulge of J1331 has an IMF \emph{more} bottom-heavy than the Salpeter IMF from fitting a NFW halo to (i) the Einstein mass and (ii) gas kinematics at larger radii $\gtrsim 8''$. In Section \ref{sec:results_JAM_NFW} we fitted a mass model with NFW halo to (i) the Einstein mass and (ii) stellar kinematics within $\sim 3.5''$. Our best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a \emph{less} bottom-heavy IMF than the Salpeter IMF. \NEW{In the following section we will argue why we do not think that the \cite{SWELLSV} model is a good model for the central regions of J1331. We therefore also do not believe their findings that the central IMF is more bottom-heavy than Salpeter.} \Wilma{[TO DO: Find new place for: "\cite{SWELLSV} found systematically lower NFW halo masses ($v_\text{circ,halo}(5'') \sim 120~\text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ,halo}(R'=5'') \sim 200~\text{km s}^{-1}$)."]}
%--------------------------------------------

@@

\NEW{In conclusion we suspect that a moderate DM contribution in the center, with some tangential anisotropy and a $\Upsilon_\text{I,*}$ which is slightly lower than a Salpeter-like IMF is the most likely model for J1331's central bulge.}

@@

\NEW{Around $R'\sim7''$ the rotation curve and dispersion show some wiggles which lead to a spontaneous rise of the $v_\text{rms}$. As the spiral arms cross the major axis around this radius, we suspect they cause this disturbance of the axisymmetric kinematics. The DM halo should also start to dominate the kinematics at larger radii, whose total mass we expect more of the order of what \Wilma{[TO DO: Dutton paper]} found---unless J1331 has a strong kinematic warp in its disk, which cannot be detected in long-slit kinematics only.}
